### 2024-08-08 paper.tex
reorganized things, started talking about jordan basis

% \par Note that \(X \mapsto A_{X, a, b}\) gives a map \(\gl_m(\mathbb{C}) \to \gl_{m + n}(\mathbb{C})\).
Let \(\mathcal{M}\) be the variety of Borel subalgebras of \(\gl_{m + n}(\mathbb{C})\).
Let
\[\widetilde{\mathcal{N}} = \{(\mathfrak{m}, a, b, X) : A_{X, a, b} \in \mathfrak{m}\} \subseteq \mathcal{M} \times \mathbb{C}^m \times \mathbb{C}^m \times \mathcal{N}.\]
Let \(\pi : \widetilde{\mathcal{N}} \to \mathcal{N}\) be the projection onto the last coordinate.
For \(X \in \mathcal{N}\), we call \(\pi^{-1}(X)\) the \emph{other Springer fiber at \(X\)}.
\section{Finding the Springer fibers of section 2}
We have \(\mathcal{M} = \{AHA^{-1} : A \in \GL_{m + n}(\mathbb{C})\}\), where \(H \subseteq \gl_{m + n}(\mathbb{C})\) is the set of upper triangular matrices.
We say a map \(X : \gl_{m + n} \to \gl_{m + n}\) preserves a flag \(V_0 \subseteq \cdots \subseteq V_{m + n}\) if \(XV_i \subseteq V_i\) for each \(i\).
Let \(E_0 \subseteq \cdots \subseteq E_{m + n}\) be the standard flag of \(\mathbb{C}^{m + n}\).
Since \(H\) is the set of \(X\) which preserve \(E\),
\[\mathcal{M} = \{\{X : \forall i. \; X(AE_i) \subseteq AE_i\} : A \in \GL_n(\mathbb{C})\}.\]
So, for \(X \in \mathcal{N}\),
\[\pi^{-1}(X) \cong \{(V, a, b) : \forall i. \; A_{X, a, b} V_i \subseteq V_i\}.\]
\par In this section we will find the irreducible components of \(\pi^{-1}(X)\).
Since \(\pi^{-1}(X) \cong \pi^{-1}(AXA^{-1})\) for any invertible \(A\), we will assume that \(X\) is in Jordan normal form.
\par Let \(\lambda\) be the shape of \(X\), and let \((e_{ij})_{i \leq r, j \leq \lambda_i}\) be a Jordan basis for \(X\), with \(X e_{ij} = e_{i,j - 1}\).
Let \(f_1, ..., f_n\) be the standard basis for \(\mathbb{C}^n\), with \(A_{X, a, b} f_i = f_{i + 1}\).

\subsection{A necessary condition for \(A_{X, a, b}\) to be nilpotent}
Suppose \(A := A_{X, a, b}\) is nilpotent.
Define the \emph{height} of a vector \(v \in \mathbb{C}^{m + n}\) as the smallest \(k \geq 0\) such that \(A_{X, a, b}^k v = 0\).
Clearly, for \(i \leq j\), we have that the height of \(f_i\) is geq the height of \(f_j\).
Thus, for each \(k \geq 0\), we must have \(A_{X, a, b}^k f_n \in \langle e_{ij} \rangle_{ij}\).
We have \(Af_n = \sum_{ij} b_{ij} e_{ij}\).
Then \(A^2 f_n = \sum_{ij} b_{ij} (e_{i, j - 1} + a_{ij} f_1)\), so we must have \(\sum_{ij} b_{ij} a_{ij} = 0\).
Similarly, since \(A^3 f_n \in \langle e_{ij}\rangle\), we see that \(\sum_{ij} b_{ij} a_{i, j - 1} = 0\).
Continuing in this way, we obtain that for all \(k \geq 0\),
\begin{equation}\label{ab_nilp}\sum_{ij} b_{ij} a_{i,j - k} = 0.\end{equation}
In fact this is in fact a sufficient condition for \(A_{X, a, b}\) to be nilpotent, as can be seen from the characteristic polynomial of \(A_{X, a, b}\).
However, we will not do this calculation here, instead just showing that it is sufficient by giving a Jordan basis.

@@

\begin{lemma}
    There is a Jordan basis \(e_{ij}\) for \(A\) such that there is at most one pair \((i, j)\) with \(f(e_{ij}) \neq 0\).
\end{lemma}
\begin{proof}
    For any change of basis \(P : V \to V\) commuting with \(A\), we obtain a new Jordan basis \((P(e_{ij}) : i \leq m, j \leq \lambda_i)\).
    For any such \(P\), define 
    \[S_P = \sum_i \begin{cases}-1, & \forall j. \; f \circ P(e_{ij}) = 0 \\ \lambda_i - \min\{j : f \circ P(e_{ij}) \neq 0\}, & \textrm{otherwise} \end{cases}.\]
    Let \(P\) be any operator, among all invertible operators commuting with \(A\), which minimizes \(S_P\).
    Write \(e_{i,j}' := P(e_{i,j})\).
    \par Suppose for contradiction that there are two distinct \(i\)'s (and some \(j\)'s) with \(f(e_{ij}') \neq 0\).
    Then we can take \(e_{i_1, j_1}'\) and \(e_{i_2, j_2}'\), where for \(k \in \{1,2\}\) we have \(f(e_{i_k, j_k}') \neq 0\), and \(\forall j < j_k. \; f(e_{i_k, j}') = 0\).
    Wlog, assume \(\lambda_{i_1} - j_1 \leq \lambda_{i_2} - j_2\).
    Then, we can define \(Q : V \to V\) by
    \begin{itemize}
        \item \(Q(e_{i_1, \lambda_1}') := e_{i_1, \lambda_1}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}'\)
        \item For \(j < \lambda_1\), \(Q(e_{i_1, j}') := A^{\lambda_{i_1} - j} Q(e_{i_1, \lambda_{i_1}}')\)
        \item For \(i \neq i_1\), \(Q(e_{ij}') = e_{ij}'\).
    \end{itemize}
    Clearly \(Q\) is invertible, and it commutes with \(A\).
    Further, I claim that \(S_{QP}  < S_P\).
    It suffices to show that \(\forall j \leq j_1. \; f \circ Q(e_{i_1,j}') = 0\).
    We have 
    \[f \circ Q(e_{i_1,j}') = f\left(A^{\lambda_{i_1} - j}\left(e_{i_1, \lambda_1}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}'\right)\right) =\]
    \[f\left(e_{i_1, j}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})} e_{i_2, j_2 + (j - j_1)}'\right) = f(e_{i_1, j}') - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})} e'_{i_2, j_2 + (j - j_1)}.\]
    Clearly (by design), this expression is zero when \(j = j_1\).
    And for \(j < j_1\), we have \(f(e'_{i_1, j}) = f(e'_{i_2, j_2 + (j_2 - j_1)}) = 0\), so it is zero then as well.
    Thus we see that \(S_{QP} < S_P\), contradicting that \(S_P\) is minimal.
    So there must be at most one \(i\) such that there exists \(j\) such that \(f(e_{ij}') \neq 0\).
    \par If there is no such \(i\), we have found the desired basis.
    So, suppose there is such an \(i_0\).
    Let \(U = \langle e_{i_0, j} \rangle\).
    Simply write \(e_j := e_{i_0, j}\).
    For any \(P : U \to U\), let \(S_P := \#\{j : f(P(e_j)) \neq 0\}\).
    Choose \(P\) to minimize \(S_P\), and then let \(e_j' := P(e_j)\).
    Suppose, for contradiction, that \(S_P > 1\).
    Then let \(j_0 = \min\{j : f(e_j') \neq 0\}\), and let \(k\) be distinct from \(j\) with \(f(e_k') \neq 0\).
    Then make the change of basis \(e_{\lambda_{i_0}}'' := e_{\lambda_{i_0}}' + ...\).
\end{proof}

\subsection{In the case of general \(X\)}
In the case where \(X\) was a single Jordan block, it was helpful to have the condition that all the nonzero \(a\)'s were to the right of all the nonzero \(b\)'s.
This allowed some tricks with right-shifting.
In the general case, the parameters \(a\) and \(b\) are not necessarily in such a helpful form.
We will see, however, that we can change basis of \(\mathbb{C}^m\) to put them in a nice form.

>>>>

% \par Note that \(X \mapsto A_{X, a, b}\) gives a map \(\gl_m(\mathbb{C}) \to \gl_{m + n}(\mathbb{C})\).
Let \(\mathcal{M}\) be the variety of Borel subalgebras of \(\gl_{m + n}(\mathbb{C})\).
Let
\[\widetilde{\mathcal{N}} = \{(\mathfrak{m}, a, b, X) : A_{X, a, b} \in \mathfrak{m}\} \subseteq \mathcal{M} \times \mathbb{C}^m \times \mathbb{C}^m \times \mathcal{N}.\]
Let \(\pi : \widetilde{\mathcal{N}} \to \mathcal{N}\) be the projection onto the last coordinate.
For \(X \in \mathcal{N}\), we call \(\pi^{-1}(X)\) the \emph{other Springer fiber at \(X\)}.

\section{A necessary condition for \(A_{X, a, b}\) to be nilpotent}
Suppose \(A := A_{X, a, b}\) is nilpotent.
Define the \emph{height} of a vector \(v \in \mathbb{C}^{m + n}\) as the smallest \(k \geq 0\) such that \(A_{X, a, b}^k v = 0\).
Clearly, for \(i \leq j\), we have that the height of \(f_i\) is geq the height of \(f_j\).
Thus, for each \(k \geq 0\), we must have \(A_{X, a, b}^k f_n \in \langle e_{ij} \rangle_{ij}\).
We have \(Af_n = \sum_{ij} b_{ij} e_{ij}\).
Then \(A^2 f_n = \sum_{ij} b_{ij} (e_{i, j - 1} + a_{ij} f_1)\), so we must have \(\sum_{ij} b_{ij} a_{ij} = 0\).
Similarly, since \(A^3 f_n \in \langle e_{ij}\rangle\), we see that \(\sum_{ij} b_{ij} a_{i, j - 1} = 0\).
Continuing in this way, we obtain that for all \(k \geq 0\),
\begin{equation}\label{ab_nilp}\sum_{ij} b_{ij} a_{i,j - k} = 0.\end{equation}
In fact this is in fact a sufficient condition for \(A_{X, a, b}\) to be nilpotent, as can be seen from the characteristic polynomial of \(A_{X, a, b}\).
However, we will not do this calculation here, instead just showing that it is sufficient by giving a Jordan basis.

\section{A Jordan basis for \(A_{X, a, b}\)}
For illustration, we will begin with the case that \(X\) is a single Jordan block.
Then we will move to the general case.

\begin{lemma}
    There is a Jordan basis \(e_{ij}\) for \(A\) such that there is at most one pair \((i, j)\) with \(f(e_{ij}) \neq 0\).
\end{lemma}
\begin{proof}
    For any change of basis \(P : V \to V\) commuting with \(A\), we obtain a new Jordan basis \((P(e_{ij}) : i \leq m, j \leq \lambda_i)\).
    For any such \(P\), define 
    \[S_P = \sum_i \begin{cases}-1, & \forall j. \; f \circ P(e_{ij}) = 0 \\ \lambda_i - \min\{j : f \circ P(e_{ij}) \neq 0\}, & \textrm{otherwise} \end{cases}.\]
    Let \(P\) be any operator, among all invertible operators commuting with \(A\), which minimizes \(S_P\).
    Write \(e_{i,j}' := P(e_{i,j})\).
    \par Suppose for contradiction that there are two distinct \(i\)'s (and some \(j\)'s) with \(f(e_{ij}') \neq 0\).
    Then we can take \(e_{i_1, j_1}'\) and \(e_{i_2, j_2}'\), where for \(k \in \{1,2\}\) we have \(f(e_{i_k, j_k}') \neq 0\), and \(\forall j < j_k. \; f(e_{i_k, j}') = 0\).
    Wlog, assume \(\lambda_{i_1} - j_1 \leq \lambda_{i_2} - j_2\).
    Then, we can define \(Q : V \to V\) by
    \begin{itemize}
        \item \(Q(e_{i_1, \lambda_1}') := e_{i_1, \lambda_1}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}'\)
        \item For \(j < \lambda_1\), \(Q(e_{i_1, j}') := A^{\lambda_{i_1} - j} Q(e_{i_1, \lambda_{i_1}}')\)
        \item For \(i \neq i_1\), \(Q(e_{ij}') = e_{ij}'\).
    \end{itemize}
    Clearly \(Q\) is invertible, and it commutes with \(A\).
    Further, I claim that \(S_{QP}  < S_P\).
    It suffices to show that \(\forall j \leq j_1. \; f \circ Q(e_{i_1,j}') = 0\).
    We have 
    \[f \circ Q(e_{i_1,j}') = f\left(A^{\lambda_{i_1} - j}\left(e_{i_1, \lambda_1}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}'\right)\right) =\]
    \[f\left(e_{i_1, j}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})} e_{i_2, j_2 + (j - j_1)}'\right) = f(e_{i_1, j}') - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})} e'_{i_2, j_2 + (j - j_1)}.\]
    Clearly (by design), this expression is zero when \(j = j_1\).
    And for \(j < j_1\), we have \(f(e'_{i_1, j}) = f(e'_{i_2, j_2 + (j_2 - j_1)}) = 0\), so it is zero then as well.
    Thus we see that \(S_{QP} < S_P\), contradicting that \(S_P\) is minimal.
    So there must be at most one \(i\) such that there exists \(j\) such that \(f(e_{ij}') \neq 0\).
    \par If there is no such \(i\), we have found the desired basis.
    So, suppose there is such an \(i_0\).
    Let \(U = \langle e_{i_0, j}' \rangle\).
    Simply write \(e_j := e_{i_0, j}'\).
    Let \(j_0 = \min\{j : f(e_j) \neq 0\}\).
    We just have to change basis to zero out \(e_j\) for \(j \neq j_0\).
    Let \(P_1(e_{\lambda_{i_0}}) := e_{\lambda_{i_0}} - \frac{f(e_{j_0 + 1})}{f(e_{j_0})} e_{\lambda_{i_0} - 1}\), and notice that \(f \circ P_1(e_j) = 0\) for all \(j \leq j_0 + 1\) except for \(j_0\).
    Then we define \(P_2(e_{\lambda_{i_0}}) := P_1(e_{\lambda_{i_0}}) - \frac{f\circ P_1(e_{j_0 + 2})}{f\circ P_1(e_{j_0})} e_{\lambda_{i_0} - 2}\), and notice that \(f \circ P_2(e_j) = 0\) for all \(j \leq j_0 + 2\) except for \(j_0\).    
    Eventually we get \(P_{\lambda_{i_0} - j_0}\), and by applying this to the \(e_j\)'s we obtain a basis of \(U\), in which there is exactly one \(j\) with \(f \circ P_{\lambda_{i_0} - j_0}(e_j) \neq 0\).
\end{proof}

\subsection{The general case}
In the case where \(X\) was a single Jordan block, it was helpful to have the condition that all the nonzero \(a\)'s were to the right of all the nonzero \(b\)'s.
This allowed some tricks with right-shifting.
In the general case, the parameters \(a\) and \(b\) are not necessarily in such a helpful form.
We will see, however, that we can change basis of \(\mathbb{C}^m\) to put them in a nice form.
\par Recall our nilpotent operator \(A_{X, a, b}\) on \(\mathbb{C}^m \times \mathbb{C}^n\).
We can define a map \(\mathbb{C}^m \to \mathbb{C}\) by \(x \mapsto x \cdot a\).
Then, by the previous lemma, we can choose a Jordan basis \(e_{ij}\) of \(\mathbb{C}^m\) so that at most one \((i, j)\) has \(f(e_{ij}) \neq 0\).
This implies that there are some \(P, k, b'\) such that \(PA_{X, a, b}P^{-1} = A_{X, 1_k, b'}\).
So, without loss of generality, we may consider this case, where \(a = 1_k\).
\par Write our basis of \(\mathbb{C}^m\) as \(e_{ij}\) and our basis of \(\mathbb{C}^n\) as \(f_j\).
Let \(i_0, r\) be such that \(a = e_{i_0, \lambda_{i_0} - r + 1}\).
Let \(b' = A^{n + r} e_{i_0, \lambda_{i_0}} = b + e_{i_0, \lambda_{i_0} - r - n}\).
Let \(b_{ij}'\) be the projection of \(b'\) onto \(e_{ij}\).
For each \(i\), let \(p_i = \max\{j : b_{ij}' \neq 0\}\).
Then for each \(i \neq i_0\), let \(q_i = \lambda_i - p_i\).  Let \(q_{i_0} = \lambda_{i_0} - r - p_{i_0}\).
Normally we index \(e_{ij}\) so that \(\lambda_i\) is nonincreasing with \(i\).
Now, it will be more convenient to assume our indices are such that \(q_i\) is nonincreasing, so we will do that.
\par We have two cases, which will closely resemble the cases we had when \(X\) was a single Jordan block.
\subsubsection{First case: \(n + \max_i p_i \geq q_{i_0} + p_{i_0}\)}
We list the chains in (roughly) decreasing order.
First, for each \(i\) such that \(q_i \geq n + r\), we take the chain of length \(p_i + q_i\) beginning with \(e_{i, p_i + q_i}\).
\par Now we handle \(e_{i_0, \lambda_{i_0}}\).
Let \(P = \max_{i : q_i < n + r} p_i\).
Set 
\[v = e_{i_0, \lambda_{i_0}} - (A^{r + n + P}e_{i_0, \lambda_{i_0}} >> r + n + P).\]
Note that \(A^{r + n} e_{i_0, \lambda_{i_0}} = b'\), so \(A^{r + n + P} e_{i_0, \lambda_{i_0}} = A^P b' = b' << P\).
Also note that if we shift \(b'\) left \(P\) times, we zero out all the rows \(i\) where \(q_i < n + r\).
This ensures that the operation of shifting \(b' << P\) right \(r + n + P\) times is invertible by applying \(A\), and thus shifting left, \(r + n + P\) times.
So, we take the chain of length \(r + n + P\) beginning with \(v\).
How do we know this is linearly independent from what came before?
Something to do with the condition on \(n\)...
\par Now, we handle the \(i\) with \(q_i < n + r\).
If \(p_i \leq \max_{k > i} p_k\), then we take the chain beginning with \(e_{i,p_i + q_i}\) of length \(p_i + q_i\).
Otherwise, we set 
\[v_i = A^{n + r - q_i} e_{i_0, \lambda_{i_0}} - \sum_{k = 1}^{p_i}b_{i,k} e_{i,k + q_i}.\]
Note that \(v_i\) is just

\section{Finding the Springer fibers of section 2}
We have \(\mathcal{M} = \{AHA^{-1} : A \in \GL_{m + n}(\mathbb{C})\}\), where \(H \subseteq \gl_{m + n}(\mathbb{C})\) is the set of upper triangular matrices.
We say a map \(X : \gl_{m + n} \to \gl_{m + n}\) preserves a flag \(V_0 \subseteq \cdots \subseteq V_{m + n}\) if \(XV_i \subseteq V_i\) for each \(i\).
Let \(E_0 \subseteq \cdots \subseteq E_{m + n}\) be the standard flag of \(\mathbb{C}^{m + n}\).
Since \(H\) is the set of \(X\) which preserve \(E\),
\[\mathcal{M} = \{\{X : \forall i. \; X(AE_i) \subseteq AE_i\} : A \in \GL_n(\mathbb{C})\}.\]
So, for \(X \in \mathcal{N}\),
\[\pi^{-1}(X) \cong \{(V, a, b) : \forall i. \; A_{X, a, b} V_i \subseteq V_i\}.\]
\par In this section we will find the irreducible components of \(\pi^{-1}(X)\).
Since \(\pi^{-1}(X) \cong \pi^{-1}(AXA^{-1})\) for any invertible \(A\), we will assume that \(X\) is in Jordan normal form.
\par Let \(\lambda\) be the shape of \(X\), and let \((e_{ij})_{i \leq r, j \leq \lambda_i}\) be a Jordan basis for \(X\), with \(X e_{ij} = e_{i,j - 1}\).
Let \(f_1, ..., f_n\) be the standard basis for \(\mathbb{C}^n\), with \(A_{X, a, b} f_i = f_{i + 1}\).

@@

\subsection{In the case of general \(X\)}
